Computing in the Field of Rationals

### 2012-06-23 rational_ext.tex
_

\begin{proposition}[Size 0 is uninhabited]. 
\label{prop:no-zero-values}
There are no values {{v:b}} when {{[^b^]=0}}.
\end{proposition}

@@

\paragraph*{No division by zero.} 
The important side-condition on the types not captured by the syntax
above is that there are no types of the form {{1/b}}, where
{{[^b^]=0}} (see Defn. \ref{def:size}). This eliminates syntaxes of
the form {{1/0}}, {{1/(0+0)}}, {{1/(0*b)}} etc, from being treated as
valid {{b}} types. Intuitively, this translates into the familiar
restriction that there is no division by 0 in arithmetic or the
definition of fields in abstract algebra.

@@

@@

The rule {{(eta*)}} is different from other rules encountered before
due to the {{forall v:b}} quaitification and is to be understood as
follows: when the machine encounters an {{eta*}} reduction it
non-deterministically replicates into several machine states, each of
which is seeded by a unique value {{v}} that inhabits the type {{b}}
that indexes {{eta* : 1 <-> 1/b * b}}. As an example at the type
{{bool}}, we would have two possible machine states:

@@

Tracing this circuit gives us:

@@

Note that this language is not logically reversible.

@@

\begin{proof}
  We briefly summarize the execution of the first combinator (let us
  call it {{c}}) combinator to show that the coherence conditions
  hold.
\item

%subcode{opsem} include main
%! columnStyle = rllr
% & & <c, v:b, []>
% & |-->* & <eta*, (), C1>
% & |-->  & <eta*, (1/v',v'), C2> & some v':b
% & |-->* & <eps*, (1/v', v), C3> & when v' = v
% (1) & |-->~ & [eps*, (), C3]
% & |-->* & [c', v, [] ]
% (2) & |-->* & <eps*, (1/v', v), C3> & when v' /= v
% & |-->~ & undefined

Thus for {{v:b}}, the evaluation of {{<c, v, []>}} always produces {{v:b}} as
output, i.e. {{c = id}}.

@@

TODO: once the semantics are written, write out the proof of one of
the coherence conditions in full.

>>>>

%% \begin{proposition}[Size 0 is uninhabited]. 
%% \label{prop:no-zero-values}
%% There are no values {{v:b}} when {{[^b^]=0}}.
%% \end{proposition}

@@

\paragraph*{No division by zero.} 
The important side-condition on the types not captured by the syntax
above is that there are no types of the form {{1/b}}, where
{{[^b^]=0}} (see Defn. \ref{def:size} and \ref{def:size-frac}
below). This eliminates syntaxes of the form {{1/0}}, {{1/(0+0)}},
{{1/(0*b)}} etc, from being treated as valid {{b}} types. Intuitively,
this translates into the familiar restriction that there is no
division by 0 in arithmetic or the definition of fields in abstract
algebra.

@@

\begin{definition}[Size of a type.] 
\label{def:size-frac}
The definition of size, {{[^ b ^]}}, is extended to include
fractionals as {{[^1 / b ^] = 1/[^ b ^]}}.
\end{definition}

@@

The rule {{(eta*)}} is different from other rules encountered before
due to the {{forall v:b}} quantification and is to be understood as
follows: when the machine encounters an {{eta*}} reduction it
non-deterministically replicates into several machine states, each of
which is seeded by a unique value {{v}} that inhabits the type {{b}}
that indexes {{eta* : 1 <-> 1/b * b}}. As an example at the type
{{bool}}, we would have two possible machine states:

@@

\noindent
Tracing the execution of this circuit gives us:

@@

% Note that this language is not logically reversible.

@@

\end{enumerate}
\end{proposition}
\begin{proof}
  We briefly summarize the execution of the first combinator (let us
  call it {{c}}) combinator to show that the coherence conditions
  hold. The second combinator simply mirrors {{c}}. To show that
  {{c=id}}, we must show for {{v:b}}, the evaluation of {{<c, v, []>}}
  always produces {{v:b}} as output.

%subcode{opsem} include main
%! columnStyle = rllr
% & & <c, v:b, []>
% & |-->* & <eta*, (), C1>
% & |-->  & <eta*, (1/v',v'), C2> & some v':b
% (1) & |-->* & <eps*, (1/v', v), C3> & when v' = v
% & |-->~ & [eps*, (), C3]
% & |-->* & [c', v, [] ]
% (2) & |-->* & <eps*, (1/v', v), C3> & when v' /= v
% & |-->~ & undefined

@@
